---
title: Spread of COVID-19
author: Ben Koos
date: '2020-05-24'
slug: spread-of-covid-19
categories: []
tags:
  - COVID-19
  - Map
type: ''
subtitle: ''
image: ''
---




If you play around with the slider on the interactive chart below you can see how COVID-19 spread throughout the United States.  


```{r echo = F}
knitr::include_app("https://ben-koos.shinyapps.io/static-covid-map/", height = "600px")
```

When I made these charts, the first thing I noticed was how much faster the disease appears to spread in coastal areas of the United States. Even in May, there are still numerous counties in the center of the country that haven't reported a single case of COVID-19.

It makes sense that states with more urban population are more connected internationally would be hit harder by the disease. Interestingly, these states tend to vote Democrat. Below you can compare the spread, total cases and cases per captia of COVID0-19 in states that voted Democrat vs. Republican in the 2016 preseidential election. 

```{r echo = F}
knitr::include_app("https://ben-koos.shinyapps.io/covid-map-by-party/", height = "1000px")
```



```{r echo = F}

require(ggplot2)
require(dplyr)
require(reshape2)

covid_data = read.csv("C:\\Users\\Ben Koos\\Documents\\Projects\\COVID 19\\Output\\COVID Data Combined.csv")


NumCases = covid_data %>% 
  group_by( date , voteD) %>% 
  summarize( cases   = sum(cases),
             tot_pop = sum(pop, na.rm = T),
             cpp     = as.character(signif(cases/tot_pop*100000,3))) %>% 
  dcast(  date ~ voteD  , value.var = c('cpp')) 

colnames(NumCases) = c("date", "repub" , "demo")

NumCases$Ratio = as.numeric(NumCases$demo)/as.numeric(NumCases$repub)

ggplot( filter(NumCases, !is.na(Ratio))) + 
  geom_line(aes(x= as.Date(date), y = Ratio), color = '#49514a', size = 1.6) +
  scale_x_date( date_labels = "%B %d, %Y" , date_breaks = "2 weeks", expand = c(0,0)) + 
  scale_y_continuous( breaks = c(0:6)) + 
  theme( axis.text.x = element_text(angle = 45, vjust = .5),
         panel.background = element_blank(),
         panel.grid.major.x = element_line(color = "#ceb69d"),
         panel.grid.major.y = element_line(color = "#ceb69d"),
         plot.title = element_text(hjust = .5)) +
  labs( y = "Ratio", x = "" , title = "Ratio of per Capita COVID-19 Cases\nStates that Voted for Clinton vs. States that Voted for Trump")
  

```

















