---
title: COVID-19 Economic Effects
author: Ben Koos
date: '2020-05-21'
slug: covid-19-economic-effects
categories:
  - R
tags:
  - COVID-19
  - FRED
  - Economics
  - Inflation
type: ''
subtitle: ''
image: ''
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

Given that I'm stuck inside becasue of COVID-19, I figured it would be a great time to start the website I had alwasy been thinking about starting. And what better post to start off with then about COVID-19. 

Given the far ranging economic consequences of forcing many millions of United States citizens to stay at home, the true economic effect of coronovirus may be inclculable. 

As someone who works in economic consulting, it's interesting to me to think that even decades from now, economists will need to control for the effects of coronoavirus in their models. 

### COVID 19: What's that got to do with the price of food?

Just take figure 1 below.  As we can see, the price of certain (but not all) types of food are rising at double digit rates (in annualized terms) beggining in April of 2020. This is far faster than in previous monthes, at least since the begging of 2019. 

It appears that the price of foods are not all affected equally. In April of 2020 the average price of meats, poultry, fish and eggs (top-right) rose at an annualized rate of more than 40%! Wheras the price index of food away from home (bottom-middle) does not appear to be effected by coronavirus at all, at least ostensibly. Other food prices most effected are  "Food and Beverages", "Food at Home"  "Cereals and Bakery Products" and "Dairy and Related Products", and "Fruits and Vegetables" but only when the price is adjusted for seasonality.  



```{r message = F}

require(reshape2)
require(ggplot2)
require(dplyr)

setwd("C:\\Users\\Ben Koos\\Documents\\Projects\\")

files = list.files("COVID 19\\FRED Data")

k = 1 
for ( file in files) { 
  data = read.csv( paste("COVID 19", "FRED Data", file,sep = "\\"))
  if(k != 1 ) { 
    Data = merge(Data, data, by = "DATE")
  } else {
    Data = data 
  }
  k = k + 1 
}
rm(list = "data")

index_name = colnames(Data)

colnames(Data)[which(index_name == "CUUR0000SEFP01")] = "Coffee"
colnames(Data)[which(index_name == "CUUR0000SAF111")] = "Cereals and Bakery Products"
colnames(Data)[which(index_name == "CUUR0000SEFJ")] = "Dairy and Related Products"
colnames(Data)[which(index_name == "CUUR0000SAF116")] = "Alcoholic Beverages"
colnames(Data)[which(index_name == "CPIFABNS")] = "Food and Beverages"
colnames(Data)[which(index_name == "CUUR0000SAF11")] = "Food at Home"
colnames(Data)[which(index_name == "CUUR0000SEFV")] = "Food Away from Home"
colnames(Data)[which(index_name == "CUUR0000SAF112")] = "Meats, Poultry, Fish, and Eggs"
colnames(Data)[which(index_name == "CPIUFDNS")] = "Food"
colnames(Data)[which(index_name == "CUUR0000SAF113")] = "Fruits and Vegetables"
colnames(Data)[which(index_name == "CWSR0000SAF113")] = "Fruits and Vegetables - Seasnonally Adjusted"

Data = melt(Data, id.vars = "DATE") %>% 
  mutate(DATE  = as.Date(DATE),
         value = as.numeric(value)) 

Deflator = Data %>% filter( DATE == as.Date("2019-1-1")) %>% 
  select(-DATE) %>% rename( deflator = value )

Data = merge(Data, Deflator, by = "variable") %>% 
  mutate( reindexed_value = value/deflator )

Data2  = Data %>%   
  arrange(variable, DATE) %>% 
  mutate( inc_rate = log(reindexed_value/lag(reindexed_value))*12) %>%
  filter(DATE >= as.Date("2019-1-1"), 
         variable != "Food",
         variable != "Food and Beverages") 

ggplot(Data2)   + 
  geom_line(aes(x = DATE, y = inc_rate, group = variable, color = variable)) +
  geom_hline( yintercept = 0) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 2)) + 
  scale_x_date( date_labels = "%b %Y", date_breaks = "2 months", expand = c(0,0)) + 
  #scale_colour_manual(values = c("#c69c8b", "#393133","#784332",
  #                               "#835b40","#7b685e","#e1c4b8",
  #                               "#a16754","#e8d4c7","#5c6469")) + 
  facet_wrap( ~variable, nrow = 3 , labeller = label_wrap_gen(width=30)) + 
  labs( y = "Monthly CPI Increase, Annualized", x = "",
        title = "Monthly Increase in Food Prices, for Urban Consumers",
        subtitle = 'January 2019 through April 2020',
        caption = "Source: FRED Consumer Food and Beverage Price Indexes") + 
  theme( panel.background = element_blank(),
         legend.position = "none", 
         axis.text = element_text(angle = 45),
         plot.title = element_text(hjust = .5),
         plot.subtitle = element_text(hjust = .5),
         panel.grid.major.x = element_line(colour="grey", linetype="dashed"))
```

